Spatial Analysis

### 2021-11-11 Scripts/spatial_analisis.Rmd
re-arranged code
Ran port analysis but realize NC has no BSB data.. weird

#Load top port data 
ports_df <- my_path("D","Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "75th") %>%
  distinct(port_name, .keep_all = T) # No need for duplication

@@

# ---------------- #
# Convert grid to sf 
# ---------------- #
grid_fp_sf <- st_as_sf(ne_grid,
             coords = c("lon", "lat"),
             crs = 4326) %>% 
  st_join(port_bf) %>% # join to port sf
  filter(!is.na(port_name)) %>%
  distinct(index,port_name, .keep_all = TRUE) #removes like 400K overlapping rows

@@

gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_fp_sf) +
    geom_sf(data = port_sf, aes(), fill = NA) +
    geom_sf(aes(color = state_postal), alpha = 0.3) +
    geom_sf(data = eez_sf, aes(),fill =NA) + 
    scale_color_manual(values = state_pallet) +
    theme_dark() +
    theme(legend.position = ""),
  # Showing each state separately
  ggplot() +
    geom_sf(data = grid_fp_sf, aes(color = state_postal),size = 0.1, alpha = 0.5) +
    facet_wrap(~state_postal) +
    theme_dark() +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet),
  nrow = 1)

@@

write_sf(grid_eez_sf_fp, my_path("R", "Partial",name="grid_eez_fp.shp"))

@@

write_csv(grid_eez_fp_df, my_path("R", "Partial/grid_fp",name = "grid_eez_fp_df.csv"))

@@

@@

# Ports data
ports_df <- my_path("D","Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "75th") %>%
  distinct(port_name, .keep_all = T) # No need for duplication

@@

#### Regulatory units map

@@

# The plot
map_plot <- land_sf %>% 
  left_join(reg_unit,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8, 
                              breaks = seq(0,55,5),
                              limits=(c(0,55))
                              )+
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

diff_plot <- land_sf %>% 
  left_join(reg_unit,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Difference (%)", alpha = 0.8,option = "C",
                              breaks = seq(0,45,5),
                              limits=(c(0,45)),
                              # labels = c("<-10",seq(-75,75,25),">100")
                              ) +
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

@@

#### Seasonal units map

@@

## Area plot

@@

## Area plot (running mean)

@@

@@

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,label,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region,spp,spatial) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
         ) %>%  
  left_join(state_lat) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State waters","Difference")
         )
  )

@@

# Plot me
ssp_area <-
  ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  geom_area(data = historic_periods,
             aes(
               x = year,
               y = n,
               color = period
             ),
             # fill = "gray90",
             alpha = 0.3,
            stat = "identity"
             ) +
  # annotate(geom= "rect",
  #          xmin = 1980, xmax = 2001, 
  #          ymin = 0, ymax = 100,
  #           fill = "gray90",
  #          colour = NA,
  #          alpha = 0.5) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State\n(Latitudinal order)",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
    scale_color_manual("",values = "grey90"
  ) +
  facet_grid(spp~spatial) +
  scale_x_continuous(breaks = seq(1979,2015,9),limits = c(1979,2015))+
  xlab("Year") +
  my_ggtheme_p(leg_pos = "right", hjust = 0.9,ax_tx_s = 14,leg_tx_s = 16, leg_tl_s = 18,ax_tl_s = 18,facet_tx_s = 16) +
  theme(legend.key.width = unit(1,"line"),
        legend.title.align=0.5)

@@

>>>>

#Load top port data 
ports_df <- my_path("D","Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "75th")

@@

# ---------------- #
# Convert grid to sf 
# ---------------- #
grid_fp_sf <- st_as_sf(ne_grid,
             coords = c("lon", "lat"),
             crs = 4326) %>% 
  st_join(port_bf) %>% # join to port sf
  filter(!is.na(port_name)) #%>%
  # distinct(index,port_name, .keep_all = TRUE) #removes like 400K overlapping rows

@@

gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_fp_sf) +
    geom_sf(data = port_sf, aes(), fill = NA) +
    geom_sf(aes(color = state_postal), alpha = 0.3) +
    geom_sf(data = eez_sf, aes(),fill =NA) + 
    scale_color_manual(values = state_pallet) +
    theme_dark() +
    theme(legend.position = ""),
  # Showing each state separately
  ggplot() +
    geom_sf(data = grid_fp_sf, aes(color = state_postal),size = 0.1, alpha = 0.5) +
    facet_wrap(~state_postal+species_name) +
    theme_dark() +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet),
  nrow = 1)

@@

write_sf(grid_eez_sf_fp, my_path("D", "Partial/grid_fp",name="grid_eez_fp.shp"))

@@

write_csv(grid_eez_fp_df, my_path("D", "Partial/grid_fp",name = "grid_eez_fp_df.csv"))

@@

# x %>% filter() %>% group_by(spp,abrev) %>% tally()

@@

# Ports data
ports_df <- my_path("D","Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "75th") %>%
  group_by(species_name,state_postal) %>% 
  summarise(n =n()) %>% 
  select(-n)
  # distinct(c(,state_postal), .keep_all = T) # No need for duplication

@@

#### F3 - Regulatory units map

@@

# The plot
map_plot <- land_sf %>% 
  left_join(reg_unit,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8, 
                              breaks = seq(0,35,5),
                              limits=(c(0,35))
                              )+
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

diff_plot <- land_sf %>% 
  left_join(reg_unit,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Difference (%)", alpha = 0.8,option = "C",
                              breaks = seq(0,25,5),
                              limits=(c(0,25)),
                              # labels = c("<-10",seq(-75,75,25),">100")
                              ) +
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

@@

#### F4 - Seasonal units map

@@

## F5 - Area plot

@@

### Area plot (running mean)

@@

#### Data preparation

```{r area_plot_Data_prep}

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  # View()
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region,spp,spatial) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
         ) %>%  
  left_join(state_lat) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State waters","Difference")
         )
  )

@@

@@

# Plot me
ssp_area <-
  ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  geom_area(data = historic_periods,
             aes(
               x = year,
               y = n,
               color = period
             ),
             # fill = "gray90",
             alpha = 0.3,
            stat = "identity"
             ) +

  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State\n(Latitudinal order)",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
    scale_color_manual("",values = "grey90"
  ) +
  facet_grid(spp~spatial) +
  scale_x_continuous(breaks = seq(1979,2015,9),limits = c(1979,2015))+
  xlab("Year") +
  my_ggtheme_p(leg_pos = "right", hjust = 0.9,ax_tx_s = 14,leg_tx_s = 16, leg_tl_s = 18,ax_tl_s = 18,facet_tx_s = 16) +
  theme(legend.key.width = unit(1,"line"),
        legend.title.align=0.5)

@@

#### Aggregated Area plot for all species
